---
title: "Chapter 6 Stats Lab"
format: html
editor: source
execute: 
  echo: true
---

## Instructions

Read through the following document, running each code chunk (and editing variables within them) as you encounter them. Use the results of the code chunks to answer the questions.

To submit your work: 1. Render the file by clicking the blue segmented arrow at the top of this window to render to a .html\
2. Submit the resulting "Chapter 6 Stats Lab.html" and "Chapter 6 Stats Lab_files" folder to the Google drive folder I shared.

## Background: Part 1

In this Lab, we will work on analyzing a provided dataset and concluding whether it seems to follow a normal distribution. To accomplish this, we will compare the empirical data to the theoretical distribution to determine if data from the experiment fits.\
The first data set, `M64_16S.csv` is data acquired from the leaderboard of Speedrun.com on completion times for the game "Super Mario 64".

::: callout-note
#### Question 1.1

Before we start, would you expect the completion times to be normally distributed?

Answer:
:::

## Getting Started

1.  Ensure `M64_16S.csv` is in the same folder as this notebook.
2.  Run the following code to load the dataset:

```{r}
# Load necessary libraries
if (!require("tidyverse")) install.packages("tidyverse")
library(readr)
library(ggplot2)
library(dplyr)

# Read dataset (assumes it's in the same directory as the .qmd file)
data <- read_csv("M64_16S.csv", show_col_types = FALSE)

# Preview the first few rows
head(data)
```

3.  Follow the guided exercises in this notebook

## Explore the data

### Histogram

Before starting any analysis, it is a good idea to explore the data.

First, we'll construct a histogram of completion times:

```{r}
# Creates histogram of time_seconds (completion time) and changes the y-axis to be density (makes area under bars add to 1)
gg_hist <- data %>%
  ggplot(aes(x = time_seconds)) + 
  geom_histogram(aes(y = after_stat(density)), binwidth = 100) # Bin width chosen to be 100s

gg_hist # display resulting plot
```

::: callout-note
#### Question 1.2

Does the data appear normally distributed?

Answer:
:::

### Summary Statistics

We can next look at the summary statistics:

```{r}
data$time_seconds %>% summary()
```

::: callout-note
#### Question 1.3

From the summary above, what is the mean completion time? What is the range of completion times?

Answer:
:::

::: callout-note
#### Question 1.4

Using the code block below, what is the standard deviation?

```{r}
data$time_seconds %>% sd()
```

Answer:
:::

::: callout-note
#### Question 1.5

Using the mean and standard deviation, what is the approximate normal distribution?

Answer: $X \sim N(___,___)$
:::

## Describe the data

### Empirical distribution (from the data)

-   What is the interquartile range? (Recall IQR = Q3 - Q1)

-   The 15th percentile is

-   The 85th percentile is

-   The median is

-   What is the theoretical probability that a randomly chosen completion time is less than 1000s?

**Code chunks to help answer prior questions:**

```{r}
x <- data$time_seconds
IQR(x)
```

```{r}
# 15th percentile
x <- data$time_seconds
prob <- .15
quantile(x, probs = prob)
```

```{r}
# 85th percentile
prob <- .5 # Change me!!
quantile(x, prob)
```

```{r}
# number of times less than 1000
less_1000 <- sum(data$time_seconds < 1000)
# total trials
total <- length(data$time_seconds)

# Probability of less than 1000?
# Recall probability of event occurring is (# of desired outcomes)/(# of total outcomes)

###########
# Fill in probability calculation here

###########
```

### Theoretical Distribution

Use the normal distribution with mean and standard deviation to answer the following:

-   What is the interquartile range?

-   The 15th percentile:

-   The 85th percentile:

-   Mean is:

-   What is the theoretical probability that a randomly chosen completion time is less than 1000s?

**Code chunks to help answer prior questions:**

```{r}
# Interquartile range
mean <- 0 # CHANGE ME!!
sd <- 1 # CHANGE ME!!

upper_quartile <- qnorm(.75, mean, sd)
lower_quartile <- qnorm(.25, mean, sd)

# IQR: 
upper_quartile - lower_quartile
```

```{r}
# 15th percentile
p <- .5 # CHANGE ME!!
mean <- 0 # CHANGE ME!!
sd <- 1 # CHANGE ME!!
qnorm(p, mean, sd)
```

```{r}
# 85th percentile
p <- .5 # CHANGE ME!!
mean <- 0 # CHANGE ME!!
sd <- 1 # CHANGE ME!!
qnorm(p, mean, sd)
```

```{r}
# Probability of less than 1000?
q <- 0 # CHANGE ME!!
mean <- 0 # CHANGE ME!!
sd <- 1 # CHANGE ME!!
pnorm(q, mean, sd)
```

## Conclusion

::: callout-note
Based on the prior analysis, do you think that the theory that the data is normally distributed holds up? Compare the calculated mean, 15/85th percentiles, probability calculation between the empirical data and theoretical distribution.

Answer:
:::

As a final look, consider the histogram from earlier with the normal distribution we analyzed on top:

```{r}
gg_hist + 
  geom_function(fun = dnorm, 
                args = list(mean = mean(data$time_seconds), 
                            sd = sd(data$time_seconds)), 
                color = "red")
```

## Background: Part 2

In this Lab, we will work on analyzing a provided dataset and concluding whether it seems to follow a normal distribution. To accomplish this, we will compare the empirical data to the theoretical distribution to determine if data from the experiment fits.\
The first data set, `SOCR-HeightWeight.csv` is data acquired from Kaggle, a website that hosts data science competitions. This is a simple dataset. It contains only the height (inches) and weights (pounds) of 25,000 different humans of 18 years of age.

::: callout-note
### Question 2.1

Before we start, would you expect the height to be normally distributed?

Answer:
:::

## Getting Started

1.  Ensure `SOCR-HeightWeight.csv` is in the same folder as this notebook.
2.  Run the following code to load the dataset:

```{r}
# Load necessary libraries
library(readr)
library(ggplot2)
library(dplyr)

# Read dataset (assumes it's in the same directory as the .qmd file)
data2 <- read_csv("SOCR-HeightWeight.csv",
                  show_col_types = FALSE) %>% 
  select("Index", "Height" = "Height(Inches)")

# Preview the first few rows
head(data2)
```

3.  Follow the guided exercises in this notebook

## Explore the data

### Histogram

Before starting any analysis, it is a good idea to explore the data.

First, we'll construct a histogram of completion times:

```{r}
# Creates histogram of time_seconds (completion time) and changes the y-axis to be density (makes area under bars add to 1)
gg_hist2 <- data2 %>%
  ggplot(aes(x = Height)) + 
  geom_histogram(aes(y = after_stat(density)), binwidth = .25) # Bin width chosen to be .25in

gg_hist2 # display resulting plot
```

::: callout-note
#### Question 2.2

Does the data appear normally distributed?

Answer:
:::

### Summary Statistics

We can next look at the summary statistics:

```{r}
data2$Height %>% summary()
```

::: callout-note
#### Question 2.3

From the summary above what is the mean height? What is the range of heights?

Answer:
:::

::: callout-note
#### Question 2.4

What is the standard deviation?

```{r}
data2$Height %>% sd()
```

Answer:
:::

::: Callout-note
#### Question 2.5

Using the mean and standard deviation, what is the approximate normal distribution?

Answer: $X \sim N(___,___)$
:::

## Describe the data

### Empirical distribution (from the data)

-   What is the interquartile range? (Recall IQR = Q3 - Q1)

-   The 15th percentile is

-   The 85th percentile is

-   The median is

-   What is the theoretical probability that a randomly chosen height is less than 6 feet (72 inches)?

**Code chunks to help answer prior questions:**

```{r}
x <- data$time_seconds
IQR(x)
```

```{r}
# 15th percentile
x <- data2$Height
prob <- .15
quantile(x, probs = prob)
```

```{r}
# 85th percentile
prob <- .5 # CHANGE ME!!
quantile(x, prob)
```

```{r}
# number of times less than 72
less_72 <- sum(data2$Height < 72)
# total trials
total <- length(data2$Height)

# Probability of less than 72?
# Recall probability of event occurring is (# of desired outcomes)/(# of total outcomes)

###########
# Fill in probability calculation here

###########
```

### Theoretical Distribution

Use the normal distribution with mean and standard deviation to answer the following:

-   What is the interquartile range?

-   The 15th percentile:

-   The 85th percentile:

-   Mean is:

-   What is the theoretical probability that a randomly chosen completion time is less than 72?

**Code chunks to help answer prior questions:**

```{r}
# Interquartile range
mean <- 0 # CHANGE ME!!
sd <- 1 # CHANGE ME!!

upper_quartile <- qnorm(.75, mean, sd)
lower_quartile <- qnorm(.25, mean, sd)

# IQR: 
upper_quartile - lower_quartile
```

```{r}
# 15th percentile
p <- .5 # CHANGE ME!!
mean <- 0 # CHANGE ME!!
sd <- 1 # CHANGE ME!!
qnorm(p, mean, sd)
```

```{r}
# 85th percentile
p <- .5 # CHANGE ME!!
mean <- 0 # CHANGE ME!!
sd <- 1 # CHANGE ME!!
qnorm(p, mean, sd)
```

```{r}
# Probability of less than 72?
q <- 0 # CHANGE ME!!
mean <- 0 # CHANGE ME!!
sd <- 1 # CHANGE ME!!
pnorm(q, mean, sd)
```

## Conclusion

::: callout-note
Based on the prior analysis, do you think that the theory that the data is normally distributed holds up? Compare the calculated mean, 15/85th percentiles, probability calculation between the empirical data and theoretical distribution.

Answer:
:::

As a final look, consider the histogram from earlier with the normal distribution we analyzed overlayed:

```{r}
gg_hist2 + 
  geom_function(fun = dnorm, 
                args = list(mean = mean(data2$Height), 
                            sd = sd(data2$Height)), 
                color = "red")
```
