---
title: "6.2 Notebook"
format: html
editor: source
---

## Calculation of Probabilities

Probabilities of normal distributions are calculated using technology.

### P(X \< x) or P(X \> x)

In R to find $P(X < x)$ (the probability of a value less than a particular number $X$) you can use the `pnorm(value, mean = 0, standard deviation = 1, lower.tail = TRUE)`.\
This function has a required argument of value, a particular x-value from the distribution we are interested in, a default mean of 0, a default standard deviation of 1, and a default to the lower tail.

To find $P(X > x)$ (the probability of a value greater than a particular number $X$), you can either do $1 - P(X < x)$ with default value of `lower.tail` or `pnorm(value, mean, standard deviation, lower.tail = FALSE)`

### Finding quantile

To find quantiles, that is the $x-value$ below which the proportion $p$ of the distribution falls, use `qnorm(probability, mean = 0, sd = 1, lower.tail = TRUE)`

### Q1

If the area to the left is 0.0228, what is the area to the right?

```{r}
left <- 0.0288
# right

```

### Q2

The final exam scores in a statistics class were normally distributed with a mean of 63 and a standard deviation of 5.

a.  Find the probability that a randomly selected student scored more than 65% on the exam

```{r}
# Pick values of p, m, sd, lower.tail appropriate to the question
# Find P(x > 65)
p <- 0 # value
m <- 0 # mean
sd <- 1 # standard deviation
lower.tail <- FALSE # below x-value, or above?
pnorm(p, m, sd, lower.tail = lower.tail)
```

b.  Find the probability that a randomly selected student scored less than 85% on the exam.

```{r}
# Pick values of p, m, sd, lower.tail appropriate to the question
# Find P(x < 85)
p <- 85 # value
m <- 63 # mean
sd <- 5 # standard deviation
lower.tail <- TRUE # below x-value, or above?
pnorm(p, m, sd, lower.tail = lower.tail)
```

c.  Find the 90th percentile, that is the x-value below which 90% of scores fall.

```{r}
# Pick values of p, m, sd, lower.tail appropriate to the question
p <- .9 # value
m <- 63 # mean
sd <- 5 # standard deviation
lower.tail <- TRUE # below x-value, or above?
qnorm(p, m, sd, lower.tail = lower.tail)
```

c.  Find the 70th percentile, that is the x-value below which 70% of scores fall.

```{r}
# Pick values of p, m, sd, lower.tail appropriate to the question
p <- .7 # value
m <- 63 # mean
sd <- 5 # standard deviation
lower.tail <- TRUE # below x-value, or above?
qnorm(p, m, sd, lower.tail = lower.tail)
```

### Q3

The golf scores for a school team were normally distributed with a mean of 68 and a standard deviation of 3. Find the probability that a randomly selected golfer scored less than 65.

```{r}
# Pick values of p, m, sd, lower.tail appropriate to the question
p <- 0 # value
m <- 0 # mean
sd <- 1 # standard deviation
lower.tail <- TRUE # below x-value, or above?
pnorm(p, m, sd, lower.tail = lower.tail)
```

### Q4

A personal computer is used for a myraid of things. Suppose the average number of hours a household personal computer is used for entertainment is 2 hours per day. Assume times for entertainment are normally distributed and the standard deviation for the times is .5 hours.

a.  Find the probability that a household computer is used for entertainment between 1.8 and 2.75 hours per day.

```{r}
# Pick values of x1, x2, mean, sd, appropriate to the question
p1 <- pnorm(x1, mean, sd) # probability less than 1.8
p2 <- pnorm(x2, mean, sd) # probability less than 2.75

# overall probability?
```

b.  Find the maximum number of hours per day that the bottom quartile of households uses a personal computer for entertainment

```{r}
# Pick values of p, mean, sd appropriate to the question
qnorm(p, mean, sd)
```

### Q5

The golf scores for a school team were normally distributed with a mean of 68 and a standard deviation of 3. Find the probability that a golfer scored between 66 and 70.

```{r}
# Find what function we need and pick appropriate variable values
```

### Q6

In the United States smartphone users between the ages of 13 and 55+ approximately follows a normal distribution with approximate mean and standard deviation of 36.9 years and 13.9 years, respectively.

a.  Determine the probability that a random smartphone user in the age range provided is between 23 and 64.7 years old.

```{r}
# Find what function we need and pick appropriate variable values

```

b.  Determine the probability that a randomly selected smartphone user in the age range provided is at most 50.8 years old.

```{r}
# Find what function we need and pick appropriate variable values
```

c.  Find the 80th percentile of this distribution.

```{r}
# Find what function we need and pick appropriate variable values
```
