---
title: "Descriptive Statistics"
format: html
---

```{r}
#| echo: false
#| message: false
#| warning: false
pkg_chk <- function(pkg) {
  if(!require(pkg, character.only = TRUE, quietly = TRUE)) install.packages(pkg, quiet = TRUE)
  
  library(pkg, character.only = TRUE)
}

pkg_chk("ggplot2")
pkg_chk("tidyverse")
```
## Useful RStudio things
Insert code chunk: `ctrl + alt + i` (Windows)  
Insert code chunk: `cmd + option + i` (MacOS)  
Run line cursor is on: `ctrl + enter` (Windows)  
Run line cursor is on: `cmd + enter` (MacOS)  
Run code chunk: `ctrl + shift + enter` (Windows)
Run code chunk: `cmd + shift + enter` (MacOS)
Render - on menu bar at top of this window, will render the document to file (html is default)


## 2.1 - Stem and Leaf, Line Graphs, Bar Graphs

### Stem and Leaf

```{r}
stem(iris$Sepal.Length)
```

### Frequency Polygon

```{r}
ggplot(data = iris, aes(x = Sepal.Length)) +
  geom_freqpoly()
```

### Bar graph

```{r}
ggplot(data = iris, aes(x = Sepal.Length)) + 
  geom_bar()
```

### Histogram

```{r}
ggplot(data = iris, aes(x = Sepal.Length)) + 
  geom_histogram()
```

## 2.3 Measures of Location in Data

### Quartiles

```{r}
summary(iris$Sepal.Length)
```
### Other quantiles
```{r}
quantile(iris$Sepal.Length, probs = .75)
```

### IQR

```{r}
IQR(iris$Sepal.Length)

quantile(iris$Sepal.Length, probs = .25)
```

### Boxplot
```{r}
library(ggplot2)
ggplot(data = iris, aes(x = Sepal.Length)) + 
  geom_boxplot(aes(color = Species))
```

### standard deviation
```{r}
sd(iris$Sepal.Length)
```

