---
title: "Introduction to R and RStudio"
author: "Nathan Young"
format: html
date: "2024-09-17"
---

## 1. Introduction to R

R is a programming language used for statistical computing and data analysis. It is widely used in academia, research, and industry. R provides various libraries and tools for data visualization, machine learning, and more.

R allows you to analyze data, create visualizations, and apply statistical techniques efficiently, which will be essential in our statistics class.

## 2. What is RStudio?

RStudio is an integrated development environment (IDE) for R. It provides a user-friendly interface to write, test, and run R code. RStudio includes several useful components that make working with R easier:

- **Console:** Where you run R commands interactively.
- **Environment Pane:** Displays variables and data loaded into memory.
- **Plots Pane:** Visualizes your data and displays graphs.
- **Script Editor:** Write and save R code for later use.

Using RStudio, you can write R scripts, test your code, and visualize results all in one place.

## 3. Installing R and RStudio

To start using R, you'll need to install both R and RStudio.

### Steps to install:

1. Download R from [The Comprehensive R Archive Network (CRAN)](https://cran.r-project.org/). Choose the appropriate version for your operating system (Windows, macOS, or Linux).
2. Download and install RStudio from [RStudio's website](https://rstudio.com/products/rstudio/download/). Once installed, open RStudio, and you'll be ready to write R code.

## 4. Basic R Commands

Let’s try some basic R commands in the RStudio Console. These commands will introduce you to simple calculations and variable assignments.

```{r}
# Basic Math
2 + 2
sqrt(16)

# Assigning variables
x <- 5
y <- 10
z <- x + y

# Print the result
print(z)
```


# 5. Working with data in R 
One of the key strengths of R is handling and analyzing data. Let’s create a small dataset (called a "vector") and perform some basic statistical calculations.

```{r}
# Create a vector of numbers
data <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)

# Calculate the mean
mean(data)

# Calculate the standard deviation
sd(data)

```

This is the foundation of working with data in R. Later, we will work with larger datasets and learn advanced techniques.

## 7. Using R for Statistics
In our statistics class, you'll need to analyze data and summarize key metrics like mean, median, and standard deviation. R makes this easy.

### Example: Summary Statistics
R includes several built-in datasets that we can use for practice. One such dataset is mtcars, which contains data about cars and their specifications.

```{r}
# Load the 'mtcars' dataset
data(mtcars)

# Get summary statistics of the dataset
summary(mtcars)
```

This command provides an overview of the dataset, showing minimum, maximum, mean, and other statistics for each variable in the dataset.