---
title: "Descriptive Statistics"
format: html
---

```{r}
#| echo: false
#| message: false
#| warning: false
pkg_chk <- function(pkg) {
  if(!require(pkg, character.only = TRUE, quietly = TRUE)) install.packages(pkg, quiet = TRUE)
  
  library(pkg, character.only = TRUE)
}

pkg_chk("ggplot2")
pkg_chk("tidyverse")
```


## 2.1 - Stem and Leaf, Line Graphs, Bar Graphs

### Stem and Leaf
To create a basic stem and leaf plot, you can use the `stem` function. 

```{r}
stem(iris$Sepal.Length)
```

### Line Graph

A line graph connects points of pairs of x and y values. 

```{r}
# The iris data doesn't lend itself well to line plots, a summary looks nicer
iris_summary <- iris %>% group_by(Sepal.Length) %>%
  summarize(n = n())
ggplot(data = iris_summary, aes(x = Sepal.Length, y = n)) + 
  geom_line()
```

### Bar Graph

```{r}
ggplot(data = iris, aes(x = Sepal.Length)) + 
  geom_bar()
```

## 2.2 Histograms, Frequency Polygons, and TIme Series Graphs

### Histogram

```{r}
ggplot(data = iris, aes(x = Sepal.Length)) + 
  geom_histogram()
```

### Frequency Polygon

```{r}
ggplot(data = iris, aes(x = Sepal.Length)) + 
  geom_freqpoly()
```

### Time Series

```{r}
EuStockMarkets2 <- as.data.frame(EuStockMarkets) %>%
  add_column("date" = as.numeric(time(EuStockMarkets)))

ggplot(data = EuStockMarkets2, aes(x = date, y = DAX)) + 
  geom_line()
```

## 2.3 Measures of location of the data

```{r}
# Will show first, second, and third quartiles along with min, max, and mean
summary(iris$Sepal.Length)
```
```{r}
# To find any quantile, change probs (as a decimal)
prob <- .25
quantile(iris$Sepal.Length, probs = prob)
```

You can find interquartile range with `IQR()`
```{r}
IQR(iris$Sepal.Length)
```

### Box plot

```{r}
ggplot(data = iris, aes(x = Sepal.Length)) + 
  geom_boxplot()
```

