---
title: "Variation activity - Dice Roll"
format: html
---

## Dice Roll

Here we will simulate the rolling of dice! The advantage of working through this on a computer is that we can roll the dice vastly many more times and quickly visualize the distribution. 

```{r}
# Define the numbers on a die
die_numbers <- c(1,2,3,4,5,6)
# How many times to roll? 
number_of_rolls <- 20
# Roll the die! Data is saved as a data frame, think of a table with column names 
# "roll_number" and "roll_value"
roll_data <- data.frame(roll_number = c(1:number_of_rolls), 
                        roll_value = sample(x = die_numbers, 
                                            size = number_of_rolls, 
                                            replace = TRUE))
```

## Investigating the data
### Looking at the data frame
Simply looking at `roll_data` doesn't give us an idea of the distribution

```{r}
# head shows first n = 20 data points
head(roll_data, 20)
```

### Using a graph
Using a graph is much more informative!

```{r}
if (!require(ggplot2)) install.packages("ggplot2")
library(ggplot2)

ggplot(data = roll_data, aes(x = roll_value)) + 
  geom_bar()
```

## What do more rolls look like? 

Increase the number of rolls a couple of times, executing the whole code block after you do so. What do you notice? 

```{r}
number_of_rolls <- 5000
# Roll the die! Data is saved as a data frame, think of a table with column names 
# "roll_number" and "roll_value"
roll_data <- data.frame(roll_number = c(1:number_of_rolls), 
                        roll_value = sample(x = die_numbers, 
                                            size = number_of_rolls, 
                                            replace = TRUE))

ggplot(data = roll_data, aes(x = roll_value)) + 
  geom_bar()
```

